Negative and Fractional Types

### 2012-06-27 rational_ext.tex

\renewcommand{\vec}[2]{\begin{bmatrix} #1 \\ #2 \end{bmatrix}}
\newcommand{\Fpx}[1]{\mathbf{F}_{#1}}
\newcommand{\modulusp}[1]{\left | #1 \right |}
\newcommand{\fname}[1]{\ulcorner #1 \urcorner}
\newcommand{\fconame}[1]{\llcorner #1 \lrcorner}

@@

In order to interpret fractional types, we first lift the values and
combinators of {{Pi}} to the world of vector spaces. Intuitively:
\begin{itemize}
\item each type {{b}} is now interpreted as a vector space of dimension
  $\modulusp{b}$ over the field $\Fpx{2}$; the elements of such a vector
  space are column vectors
\item each value {{v_i : b}} is interpreted as column vector:
\[\begin{array}{r@{~}l}
\begin{array}{r} 
v_0 \\ v_1 \\ \vdots \\ v_i \\ \vdots \\ v_n 
\end{array} & 
\begin{bmatrix}
0 \\ 0 \\ \vdots \\ 1 \\ \vdots \\ 0
\end{bmatrix}
\end{array}\]
\item each combinator {{c : b1 <-> b2}} is interpreted as a matrix whose
  columns are indexed by the values of type {{b1}} and whose rows are indexed
  by the values of type {{b2}}. The matrix is \emph{orthogonal} with each row
  and column containing exactly one occurrence of the value 1.
\end{itemize}

Once we are in the world of vector spaces, the fractional type {{1/b}} is
interpreted as the \emph{dual} to the vector space denoted by {{b}}. The
elements of this dual vector space can be represented as \emph{row vectors}.

Multiplying a row vector by a column gives a scalar: this is eta and epsilon.

>>>>

\newcommand{\vecspace}[1]{\mathbb{V}_{#1}}
\renewcommand{\vec}[2]{\begin{bmatrix} #1 \\ #2 \end{bmatrix}}
\newcommand{\Fpx}[1]{\mathbb{F}_{#1}}
\newcommand{\modulusp}[1]{\left | #1 \right |}
\newcommand{\fname}[1]{\ulcorner #1 \urcorner}
\newcommand{\fconame}[1]{\llcorner #1 \lrcorner}

@@

In order to interpret fractional types, we first lift the values and
combinators of {{Pi}} to the world of vector spaces. Intuitively:
\begin{itemize}
\item instead of interpreting a type {{b}} as a set of cardinality
  $\modulusp{b}$, we now interpret each type {{b}} as a vector space
  $\vecspace{b}$ of dimension $\modulusp{b}$ over the field of booleans
  $\Fpx{2}$;
\item the elements of such a vector space are column vectors whose elements
  are 0 or 1 and which indexed by the values of type {{b}}, i.e., 
  a value {{v_i : b}} is interpreted as the vector:
\[\begin{array}{r@{~}l}
\begin{array}{r} 
v_0 \\ v_1 \\ \vdots \\ v_i \\ \vdots \\ v_n 
\end{array} & 
\begin{bmatrix}
0 \\ 0 \\ \vdots \\ 1 \\ \vdots \\ 0
\end{bmatrix}
\end{array}\]
\item each combinator {{c : b1 <-> b2}} is interpreted as a matrix whose
  columns are indexed by the values of type {{b1}} and whose rows are indexed
  by the values of type {{b2}}. The matrix is \emph{orthogonal} with each row
  and column containing exactly one occurrence of the value 1.
\end{itemize}

Now that the types are interpreted as vector spaces, it is possible to add a
new type {{1/b}} which corresponds to the \emph{dual} of the vector space
denoting {{b}}. The elements of this dual vector space are \emph{linear
  functionals} mapping vectors in $V_b$ to a scalar (0 or 1). The isomorphism
between the types {{1}} and {{b * 1/b}} is now a consequence of that fact
that the pairing of a functional in the dual space $\vecspace{1/b}$ and a
vector in the space $\vecspace{b}$ produces a scalar.

The alert co-author should have noticed that using $\Fpx{2}$ as the underlying
field is actually nonsense. To get things to work properly we need a field of
one element, which apparently exists and has some nice properties.
